Effects of Unemployment Benefits and Uncertainty in Heterogeneous Models

### 2017-01-24 presentation/presentation.tex
Include reference

\documentclass{beamer}
\usepackage{graphicx}
\usepackage{breqn}
\graphicspath{ {images/Eric/} }
\usepackage{amsmath}
\usepackage{dsfont}
% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

@@

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Changes in idiosyncratic risk}

\begin{frame}{Questions}
	\begin{itemize}

	\item {
	Does idiosyncratic risk matter? How do changes in such risk affect other economic variables? Are there distributional consequences?
	}

	\end{itemize}
\end{frame}

\begin{frame}{How do we examine these questions?}
	\begin{itemize}

	\item {
	We will look at changes in the unemployment benefit. 
	}

	\end{itemize}
\end{frame}

\subsection{The model}
\begin{frame}{Model Summary}
	\begin{itemize}

	\item {
	Ex ante identical individuals are subject to uninsurable idiosyncratic shocks. 
	}
	\item {
Agents can interact on the asset market.
	}
	\item {
The model generates an endogenous distribution of wealth across consumers. 
	}	
	\item {
Aggregates and prices are jointly determined by the saving behavior of the collection of all households. 
}

	\end{itemize}

\end{frame}

\subsection{Changes in the unemployment insurance}

\begin{frame}{Different parameters of the model}

  \centering{\includegraphics[scale=0.45]{parameters}}

\end{frame}

\begin{frame}{Different levels of risk aversion}

  \centering{\includegraphics[scale=0.45]{parameters_sigma}}

\end{frame}

\begin{frame}{First Observations}
	\begin{itemize}

	\item {
	Changes in idiosyncratic risk do affect aggregate variables.
	}

	\item {
	The magnitude of impact varies for different levels of risk aversion.
	}

		\item {
	How do the different levels of the aggregates compare to a situation with perfect insurance? 
	}

	\end{itemize}

\end{frame}

\begin{frame}{Comparing to complete markets}
  \centering{\includegraphics[scale=0.45]{log-deviation}}

\end{frame}

\begin{frame}{Comparing to complete markets}
	\begin{itemize}

	\item {
	As the insurance increases, the aggregates converge to the levels of the complete market economy. 
	}

	\end{itemize} 
\end{frame}

\begin{frame}{Distribution of capital}

\begin{figure}[!tbp]
  \centering
  \begin{minipage}[b]{0.325\textwidth}
    \includegraphics[width=\textwidth]{distribution1}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.325\textwidth}
    \includegraphics[width=\textwidth]{distribution2}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.325\textwidth}
    \includegraphics[width=\textwidth]{distribution3}
  \end{minipage}
\end{figure}
\ \ \ \ \ \ \  rw = 0.01 \ \ \ \ \ \ \ \ \ \ \ \ rw = 0.42579 \ \ \ \ \ \ \ \ \ \ \ \ \ rw = 0.8
\end{frame}

\begin{frame}{Concluding on the distribution}

\begin{itemize}

	\item {
As precautionary savings decrease, aggregate capital decreases, thus increasing the interest rate and reducing the wage. 
}
	\item {
The increase in unemployment benefits further reduces the net wage of the employed. The government has to increase income taxes to finance the higher unemployment benefits.
}
	\item {
Therefore, with higher unemployment benefits the agent's income depends more on his accumulated capital than his labor or benefits income. 
}
	\item {
The dispersion of the distribution therefore increases as benefits increase.
}

\end{itemize}

\end{frame}

\begin{frame}{Concluding the first part}

	\begin{itemize}

	\item {
Idiosyncratic risk matters: 
Precautionary savings increase with higher uncertainty, thus increasing aggregate capital. 
	}	
	\item {
Increasing unemployment benefits has distributional consequences. 
	}

	\end{itemize} 
\end{frame}

\begin{frame}{The limitations}

	\begin{itemize}

	\item {The model's labor supply is exogenous. Thus, changing the unemployment benefits has no impact on labor supply or the transition probabilities.
}	
	\item {This contradicts empirical evidence: Decker(1997).
	}
	\item{Endogenous labor matters: Mukoyama(2012).
	}

	\end{itemize} 
\end{frame}

\section{Welfare effects}
\subsection{}

\section{Aggregate risk}
\subsection{}

\section{Conclusions}
\subsection{}

>>>>

\documentclass{beamer}
\usepackage{graphicx}
\usepackage{breqn}
\graphicspath{ {images/} }
\usepackage{amsmath}
\usepackage{dsfont}
% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

@@

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Part I}
\subsection{}
\begin{frame}{Introduction}
  \begin{itemize}

  \item {1.}
  \end{itemize}
\end{frame}
\section{Part II}
\subsection{}
\begin{frame}{Welfare Analysis}
  \begin{itemize}
  \item {
  Which variable should we look at if we want to say consumers are better off? Output? Consumption?
  }
  \item {
  We want to measure somehow whether consumer welfare increased or not.
  }
  \item {
  But how to measure welfare? Can we compare the utility levels of consumers? And whose utility should we compare in a heterogeneous agent framework?
  }
  \end{itemize}
\end{frame}

\begin{frame}{Life-time Utility}
  \begin{itemize}
  \item {
  We calculate the life-time utility of each agent when the economy is in steady state. One time before the policy change and once after the policy change.
  }
  \item {
  Cannot compare two steady states directly. We need to take the transition into account.
  }
  \item {
  Calculating the exact transition path is troublesome. We follow the approximation by \cite{KMS2010} which takes care of the idiosyncratic transition.
  }
  	\begin{enumerate}
  	\item {
  	Just place the agents in the first economy with their current assets into the second economy.
  	}
  	\end{enumerate}
  \end{itemize}
\end{frame}

\begin{frame}{Welfare Analysis}
  \begin{itemize}
  \item {
  Now we can compare the utility level of the same agent before and after the policy change.
  }
  \item {
  Need to come up with a way to quantify the utility differences sensibly.
  }
  \item {
  We calculate a consumption and a cash equivalent of the policy change.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Consumption Equivalent}
  \begin{itemize}

  \item {
  Permanent change of consumption by factor $\Delta$ that increases life-time
  utility by same amount as policy change:
  }
  \end{itemize}
  \begin{equation}
  U^{2}(e,k) = \mathbb{E}\sum_{t=0}^{\infty}\beta^{t}\frac{(c^{1}_{i,t}\Delta)^{1-\sigma}-1}{1-\sigma}\nonumber
  \end{equation}
  \begin{itemize}
  \item {
  If $ \Delta>1 $ agents prefer policy change $ (U^{2}) $, otherwise they do not want it.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Consumption Equivalent}
  \centering{\includegraphics[scale=0.6]{cons_equiv}}
\end{frame}

\begin{frame}{Consumption Equivalent}
  \centering{\includegraphics[scale=0.6]{cons_equiv_unemp_emp}}
\end{frame}

\begin{frame}{Consumption Equivalent vs Wealth}
  \centering{\includegraphics[scale=0.6]{cons_equiv_wealth}}
\end{frame}

\begin{frame}{Consumption Equivalent}
\framesubtitle{Summary}
  \begin{itemize}
  \item {
  There exists a welfare increase in the mean and in the median.
  }
  \item {
  However, there is disagreement on whether to increase or decrease the unemployment benefits.
  }
  \item {
  Careful: the consumption equivalent is biased towards poorer agents as relative increases for them are higher than for already high-consumption agents.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Cash Equivalent}
  \begin{itemize}
  \item {
  Transfer of $\Delta$ units of wealth in the first period
  }
  \end{itemize}

  \begin{equation}
  U^{2}(e,k) = U^{1}(e,k+\Delta) \nonumber
  \end{equation}

  \begin{itemize}
  \item {
  If $ \Delta>0 $ agents prefer the policy change $ (U^{2}) $, otherwise not.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Cash Equivalent}
  \centering{\includegraphics[scale=0.6]{cash_equiv}}
\end{frame}

\begin{frame}{Cash Equivalent}
  \centering{\includegraphics[scale=0.6]{cash_equiv_unemp_emp}}
\end{frame}

\begin{frame}{Cash Equivalent}
\framesubtitle{Summary}
  \begin{itemize}
  \item {
  There does not exist a higher welfare state in the mean, only in the median.
  }
  \item {
  If a higher welfare state existed for the mean we could redistribute wealth such that everyone would be better off. \\
  $\rightarrow$ Pareto improvement.
  }
  \item {
  One can clearly see the friction between unemployed and employed.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Cash vs. Consumption Equivalent}
\centering{\includegraphics[scale=0.6]{cash_cons_equiv}}
\end{frame}

\begin{frame}{Summary}
  \begin{itemize}
  \item {
  Agents do not want full unemployment insurance although there is no moral hazard.
  }
  \item {
  Different groups in society want different policies. We did not find a policy reform that could make everyone better off with certainty.
  }
  \item {
  Of course, the model is not complete. A next step would be to endogenize the unemployment rate such that it reacts to changes in unemployment benefits.
  }
  \end{itemize}
\end{frame}

\section{Part III}
\subsection{}
